---
title: Add resources
description: Accessing additional resources inside a Viash component.
order: 20
---

{{< include ../../_includes/_language_chooser.qmd >}}

```{r setup, include=FALSE}
repo_path <- system("git rev-parse --show-toplevel", intern = TRUE)
source(paste0(repo_path, "/_includes/_r_helper.R"))
source(paste0(repo_path, "/guide/component/_language_examples.R"))
```

If a script needs access to an external file, it needs to be added as a resource in the config.


```{r setup-adding-resources, include=FALSE}
temp_dir <- tempfile("add-dependencies")
dir.create(temp_dir, recursive = TRUE, showWarnings = FALSE)
on.exit(unlink(temp_dir, recursive = TRUE), add = TRUE)

# create tempdir with modified files
add_text_resource <- function(config) {
  config$functionality$resources <- c(
    config$functionality$resources, 
    list(list(path = "my_file.txt"))
  )
  config
}
add_line_after_viash_end <- function(...) {
  function(script) {
    which_end <- grep("VIASH END", script)
    c(
      script[seq_len(which_end)],
      "",
      ...,
      #"",
      script[-seq_len(which_end)]
    )    
  }
}
use_text_resource <- list(
  bash = add_line_after_viash_end(
    "# view resource file", 
    "cat \"$meta_resources_dir/my_file.txt\""
  ),
  csharp = add_line_after_viash_end(
    "// view resource file",
    "string myFile = $\"{meta.resources_dir}/my_file.txt\";",
    "string text = File.ReadAllText(myFile);",
    "Console.WriteLine(text);"
  ),
  js = add_line_after_viash_end(
    "// view resource file",
    "const my_file = `${meta['resources_dir']}/my_file.txt`",
    "fs.readFile(my_file, 'utf8', (err, data) => {",
    "  if (err) throw err;",
    "  console.log(data);",
    "});"
  ),
  python = add_line_after_viash_end(
    "# view resource file",
    "my_file = f\"{meta['resources_dir']}/my_file.txt\"",
    "with open(my_file, \"r\") as f:",
    "  print(f.read())"
  ),
  r = add_line_after_viash_end(
    "# view resource file",
    "lines <- readLines(paste0(meta$resources_dir, \"/my_file.txt\"))",
    "cat(lines, sep = \"\\n\")"
  ),
  scala = add_line_after_viash_end(
    "// view resource file",
    "val myFile = s\"${meta.resources_dir}/my_file.txt\"",
    "val src = scala.io.Source.fromFile(myFile)",
    "src.getLines.foreach(println)"
  )
)
langs <- langs %>%
  mutate(
    label = gsub("#", "\\\\#", label),
    config_path = paste0(temp_dir, "/", id, "/", basename(example_config)),
    script_path = paste0(temp_dir, "/", id, "/", basename(example_script))
  )
pwalk(langs, function(id, label, example_config, example_script, config_path, script_path, ...) {
  # create dir
  dir.create(paste0(temp_dir, "/", id), recursive = TRUE, showWarnings = FALSE)

  # todo: create below
  readr::write_lines("contents of my_file.txt", paste0(temp_dir, "/", id, "/my_file.txt"))

  # read config, make modifications, write to file
  conf <- yaml::read_yaml(example_config) %>%
    add_text_resource()
  write_yaml(
    conf,
    config_path
  )
  
  # modify script and write to file
  script <- readr::read_lines(example_script) %>%
    use_text_resource[[id]]()
  readr::write_lines(script, script_path)
    
})
```

First, create a file called `my_file.txt`.

::: {.panel-tabset}
```{r create-file, output="asis"}
pwalk(langs, function(id, label, config_path, script_path, ...) {
  qrt(
    "## {% label %}
    |
    |```{sh}
    |echo 'Here are the contents of my_file.txt' > my_file.txt
    |```
    |
    |", .dir = paste0(temp_dir, "/", id))
})
```
:::

Next, the file needs to be added to the config as a resource.
This will let Viash know to copy the new file inside a component's resource directory.

::: {.panel-tabset}
```{r show-config, output="asis"}
pwalk(langs, function(id, label, config_path, script_path, ...) {
  qrt(
    "## {% label %}
    |
    |```yaml
    |{% paste(readr::read_lines(config_path), collapse = '\n    |') %}
    |```
    |")
})
```
:::

Lastly, to access a resource from within the script, 
use the `resources_dir` [meta-variable](variables.qmd):

::: {.panel-tabset}
```{r show-script, output="asis"}
pwalk(langs, function(id, label, config_path, script_path, ...) {
  qrt(
    "## {% label %}
    |
    |```{% id %}
    |{% paste(readr::read_lines(script_path), collapse = '\n') %}
    |```
    |
    |")
})
```
:::

Now we can run the component as follows:

::: {.panel-tabset}
```{r run-component, output="asis"}
pwalk(langs, function(id, label, config_path, script_path, ...) {
  qrt(
    "## {% label %}
    |
    |```{bash}
    |viash run config.vsh.yaml -- --input config.vsh.yaml --output foo.txt
    |```
    |
    |", .dir = paste0(temp_dir, "/", id))
})
```
:::

{{< include ../../_includes/_prune_all_images.qmd >}}