---
title: Unit testing
description: How to ensure high quality components.
search: true
engine: knitr
order: 50
---


{{< include ../../includes/_language_chooser.qmd >}}

```{r setup, include=FALSE}
repo_path <- system("git rev-parse --show-toplevel", intern = TRUE)
source(paste0(repo_path, "/includes/_r_helper.R"))
source(paste0(repo_path, "/guide/component/_language_examples.R"))

langs <- langs %>%
  mutate(
    label = gsub("#", "\\\\#", label)
  )
```

We recommend adding unit tests to all of your components in a project as soon as possible.
Implementing unit tests not only ensures that your component works, but it also makes
the project more maintainable in the long run as you will be notified when a component breaks.

This page describes how to add a unit test to your component.

```{r setup-adding-resources, include=FALSE}
temp_dir <- tempfile("unit-testing")
dir.create(temp_dir, recursive = TRUE, showWarnings = FALSE)
on.exit(unlink(temp_dir, recursive = TRUE), add = TRUE)

# create tempdir with modified files
add_unit_test <- function(config, type, path) {
  # add helper to resources
  config$functionality$test_resources <- list(
    list(type = type, path = path)
  )
  config
}
langs <- langs %>%
  mutate(
    example_test = gsub("script", "test", example_script),
    config_path = paste0(temp_dir, "/", id, "/", basename(example_config)),
    script_path = paste0(temp_dir, "/", id, "/", basename(example_script)),
    test_path = paste0(temp_dir, "/", id, "/", basename(example_test)),
  )
pwalk(langs, function(id, script_type, example_config, example_script, example_test, config_path, script_path, test_path, ...) {
  # create dir
  dir.create(paste0(temp_dir, "/", id), recursive = TRUE, showWarnings = FALSE)

  # read config, make modifications, write to file
  conf <- yaml::read_yaml(example_config) %>%
    add_unit_test(script_type, basename(test_path))
  write_yaml(conf, config_path)
  
  # copy script
  file.copy(example_script, script_path)
  file.copy(example_test, test_path)
})
```


## Create unit test

Below is an example of how to add a unit test to a Viash component.

::: {.panel-tabset}
```{r show-helper, output="asis"}
pwalk(langs, function(id, label, test_path, ...) {
  qrt(
    "## {% label %}
    |
    |```{% id %}
    |{% paste(readr::read_lines(test_path), collapse = '\n') %}
    |```
    |
    |", .dir = paste0(temp_dir, "/", id))
})
```
:::

1. Create an input file "foo.txt" with some sample content.
2. Run an executable file specified in the "meta" dictionary with the arguments "--input foo.txt" and "--output bar.txt".
3. Check whether the output file "bar.txt" exists.
4. Check whether the content of the output file is the same as the content of the input file.
5. If both checks pass, the script prints "Test finished successfully".

:::{.callout-important}
A test script doesn't need to be written in the same scripting language as the main script, as long as all of the required dependencies are available. This means the main script could be written in R, but the unit test could be written in Bash.
:::

## Add test to config

Next, we need to add the unit test to the `test_resources` section in the Viash config.

::: {.panel-tabset}
```{r show-config, output="asis"}
pwalk(langs, function(id, label, config_path, script_path, ...) {
  qrt(
    "## {% label %}
    |
    |```yaml
    |{% paste(readr::read_lines(config_path), collapse = '\n    |') %}
    |```
    |", .dir = paste0(temp_dir, "/", id))
})
```
:::

Note that if you can add multiple unit tests to the `test_resources` section,
they will all be evaluated. You can also add other resources, similar to 
what is described in the page on [adding resources](add-resources.qmd).

## Run the unit test

You can run the unit test as follows:

::: {.panel-tabset}
```{r echo=FALSE, output="asis"}

pwalk(langs, function(id, label, config_path, script_path, ...) {
  qrt(
    "## {% label %}
    |
    |```{bash test-example}
    |viash test config.vsh.yaml
    |```
    |", 
    .dir = paste0(temp_dir, "/", id)
  )
})
```
:::

When running `viash test`, Viash will follow the following steps:

1. Create a temporary directory
2. Build the component into the main executable
3. (Re-)build the Docker image for the component
4. Iterate over all unit test scripts:
  a. Build the unit test into an executable
  b. Run the unit test, passing the main executable as an argument
5. Return exit code 0 if all of the above steps succeed, otherwise 0

## Bonus: unit testing all of the components

If you have multiple Viash components located in a directory
called `src/`, what happens when you run the following?

```bash
viash ns test --parallel --src src/
```
