---
title: Add dependencies
search: true
engine: knitr
---

{{< include ../../includes/_language_chooser.qmd >}}

```{r setup, include=FALSE}
repo_path <- system("git rev-parse --show-toplevel", intern = TRUE)
source(paste0(repo_path, "/includes/_r_helper.R"))
source(paste0(repo_path, "/guide_v2/component/_language_examples.R"))

temp_dir <- tempfile("add-dependencies")
dir.create(temp_dir, recursive = TRUE, showWarnings = FALSE)
on.exit(unlink(temp_dir, recursive = TRUE), add = TRUE)

# create tempdir with modified files
add_setup <- function(...) {
  function(config) {
    config$functionality$name <-
      paste0(config$functionality$name, "_with_setup")
    config$platforms <- map(config$platforms, function(platform) {
      if (platform$type == "docker") {
        platform$setup <- list(...)
      }
      platform
    })
    config
  }
}
config_overrides <- list(
  bash = add_setup(list(type = "apk", packages = c("curl", "wget"))),
  csharp = add_setup(list(type = "apk", packages = c("curl", "wget"))),
  js = add_setup(list(type = "apt", packages = c("curl", "wget"))),
  python = add_setup(
    list(type = "apt", packages = c("curl", "wget")),
    list(type = "python", packages = "anndata")
  ),
  r = add_setup(
    list(type = "apt", packages = c("curl", "wget")),
    list(type = "r", packages = "tidyverse")
  ),
  scala = add_setup(list(type = "apt", packages = c("curl", "wget")))
)
langs <- langs %>%
  mutate(
    label = gsub("#", "\\\\#", label),
    config_path = paste0(temp_dir, "/", id, "/", basename(example_config)),
    script_path = paste0(temp_dir, "/", id, "/", basename(example_script))
  )
pwalk(langs, function(id, label, example_config, example_script, config_path, script_path, ...) {
  # create dir
  dir.create(paste0(temp_dir, "/", id), recursive = TRUE, showWarnings = FALSE)
  
  # copy script
  file.copy(example_script, script_path)

  # read config, make modifications, write to file
  conf <- yaml::read_yaml(example_config)
  conf2 <- config_overrides[[id]](conf)
  handlers <- list(
    logical = function(x) {
      result <- ifelse(x, "true", "false")
      class(result) <- "verbatim"
      return(result)
    }
  )
  yaml::write_yaml(
    conf2,
    config_path,
    handlers = handlers,
    indent.mapping.sequence = TRUE
  )
})
```

In the previous section, reproducibility of our Viash component was ensured by a predefined Docker image such as `bash:4.0` and `python:3.10`. However, your script might require other software dependencies, such as command-line tools or Python and R packages.

By default, Viash will build component-specific Docker images. This means that every Viash component can have its own set of dependencies.

## Extended example

Below is an example where additional software is added to a base Docker image using the [`setup`]({{< var reference.setup >}}) section of a Docker platform.

::: {.panel-tabset}

```{r create-config, output="asis"}
pwalk(langs, function(id, label, config_path, ...) {
  qrt(
    "## {% label %}
    |
    |
    |```yaml
    |{% paste(readr::read_lines(config_path), collapse = '\n    |') %}
    |```
    |
    |")
})
```
:::


You can (re)build a component's Docker image by passing the `---setup` flag to the executable:

::: {.panel-tabset}
```{r echo=FALSE, output="asis"}

pwalk(langs, function(id, label, config_path, script_path, ...) {
  qrt(
    "## {% label %}
    |
    |Build the executable:
    |
    |```{bash}
    |viash build config.vsh.yaml --platform docker --output target
    |```
    |
    |Build the Docker image:
    |
    |```{bash}
    |target/example_{% id %}_with_setup ---setup cachedbuild
    |```
    |", 
    .dir = paste0(temp_dir, "/", id)
  )
})
```
:::

Alternatively, you can also build the executable _and_ it's corresponding Docker image in one go:

::: {.panel-tabset}
```{r echo=FALSE, output="asis"}

pwalk(langs, function(id, label, config_path, script_path, ...) {
  qrt(
    "## {% label %}
    |
    |```{bash}
    |viash build config.vsh.yaml --platform docker --output target --setup cachedbuild
    |```
    |", 
    .dir = paste0(temp_dir, "/", id)
  )
})
```
:::

## Steps for creating a custom Docker platform

Here is a series of steps you can follow to add a Docker platform to your Viash component from scratch.

### Step 1: Choose a base image

To start off, you'll need to choose a base Docker image to start working with. In deciding which base image to use, it's important to consider the size of the image and how trustworthy the source image is.

:::{.callout-tip}
If the container does not have Bash installed, don't forget to install this in Step 2.
:::

Here is a list of base images we commonly use:

* Bash: [`bash`](https://hub.docker.com/_/bash), [`ubuntu`](https://hub.docker.com/_/ubuntu)
* C#: [`ghcr.io/data-intuitive/dotnet-script`](https://github.com/data-intuitive/ghcr-dotnet-script/pkgs/container/dotnet-script)
* JavaScript: [`node`](https://hub.docker.com/_/node)
* Python: [`python`](https://hub.docker.com/_/python), [`nvcr.io/nvidia/pytorch`](https://catalog.ngc.nvidia.com/orgs/nvidia/containers/pytorch)
* R: [`eddelbuettel/r2u`](https://hub.docker.com/r/eddelbuettel/r2u), [`rocker/tidyverse`](https://hub.docker.com/r/rocker/tidyverse)
* Scala: [`sbtscala/scala-sbt`](https://hub.docker.com/r/sbtscala/scala-sbt)

Feel free to suggest other important base images by sending us a pull request on GitHub.



### Step 2: Installing additional dependencies

You can use the `setup` section to many different types of layers. Here are some examples:

- Apk requirements:
  ```yaml
  setup:
    - type: apk
      packages: [ curl ]
  ```
- Apt requirements:
  ```yaml
  setup:
    - type: apk
      packages: [ curl ]
  ```
- Docker requirements:
  ```yaml
  setup:
    - type: docker
      build_args: "R_VERSION=hello_world"
      run: |
        echo 'Run a custom command'
        echo 'Foo' > /path/to/file.txt
  ```
- Javascript requirements:
  ```yaml
  setup:
    - type: javascript
      packages: [ express ]
      github: [ "expressjs/express" ]
  ```
- Python requirements:
  ```yaml
  setup:
    - type: python
      packages: [ anndata ]
    github: [ jkbr/httpie ]
  ```
- R requirements:
  ```yaml
  setup:
    - type: r
      packages: [ anndata ]
      bioc: [ AnnotationDbi, SingleCellExperiment ]
      github: rcannood/SCORPIUS
  ```
- Ruby requirements:
  ```yaml
  setup:
    - type: ruby
      packages: [ pry ]
  ```
- Yum requirements:
  ```yaml
  setup:
    - type: ruby
      packages: [ pry ]
      github: [ "pry/pry" ]
  ```

For more information on the possible setup entries, check out the [reference documentation]({{< var reference.setup >}}).

:::{.callout-important}
Don't forget to rebuild the Docker image after making changes to the `setup` section of your Docker platform (see next step).
:::

### Step 3: Rebuild Docker image

After adding additional setup entries, it's important to **rerun** ```---setup cachedbuild` to rebuild the Docker image, as Viash will not rebuild the Docker image when it already exists.

::: {.panel-tabset}
```{r echo=FALSE, output="asis"}

pwalk(langs, function(id, label, config_path, script_path, ...) {
  qrt(
    "## {% label %}
    |
    |```{bash}
    |viash build config.vsh.yaml --platform docker --output target --setup cachedbuild
    |```
    |", 
    .dir = paste0(temp_dir, "/", id)
  )
})
```
:::

You can choose what strategy to build an executable with when using a Docker backend by passing the `--setup` option followed by one of the strategies below.  


Building an image:

* `alwaysbuild` / `build` / `b`: Always build the image from the dockerfile. This is the default setup strategy.
* `alwayscachedbuild` / `cachedbuild` / `cb`: Always build the image from the dockerfile, with caching enabled.
* `ifneedbebuild`: Build the image if it does not exist locally.
* `ifneedbecachedbuild`: Build the image with caching enabled if it does not exist locally.

Pulling an image:

* `alwayspull` / `pull` / `p`: Try to pull the container from [Docker Hub](https://hub.docker.com) or the [specified docker registry](/reference/config/platforms/DockerPlatform.html#registry).
* `alwayspullelsebuild` / `pullelsebuild`: Try to pull the image from a registry and build it if it doesn't exist.
* `alwayspullelsecachedbuild` / `pullelsecachedbuild`: Try to pull the image from a registry and build it with caching if it doesn't exist.
* `ifneedbepull`: If the image does not exist locally, pull the image.
* `ifneedbepullelsebuild` If the image does not exist locally, pull the image. If the image does exist, build it.
* `ifneedbepullelsecachedbuild`: If the image does not exist locally, pull the image. If the image does exist, build it with caching enabled.

Pushing an image:

* `push`: Push the container to [Docker Hub](https://hub.docker.com)  or the [specified docker registry](/reference/config/platforms/DockerPlatform.html#registry).
* `pushifnotpresent` Push the container to [Docker Hub](https://hub.docker.com) or the [specified docker registry](/reference/config/platforms/DockerPlatform.html#registry) if the [specified tag](/reference/config/platforms/DockerPlatform.html#tag) does not exist yet.

Doing nothing:

* `donothing` / `meh`: Do not build or pull anything.

## Troubleshooting

Below are several steps that might help you troubleshoot the image when the setup fails.

### View Dockerfile

You can view the actual Dockerfile used by Viash by passing the `---dockerfile` flag:

::: {.panel-tabset}
```{r echo=FALSE, output="asis"}

pwalk(langs, function(id, label, config_path, script_path, ...) {
  qrt(
    "## {% label %}
    |
    |```{bash}
    |target/example_{% id %}_with_setup ---dockerfile
    |```
    |", 
    .dir = paste0(temp_dir, "/", id)
  )
})
```
:::

### Enter debugging session

You can also hop in a Bash session inside the Docker image using the `---debug` flag:

::: {.panel-tabset}
```{r echo=FALSE, output="asis"}

pwalk(langs, function(id, label, config_path, script_path, ...) {
  qrt(
    "## {% label %}
    |
    |```bash
    |target/example_{% id %}_with_setup ---debug
    |```
    |
    |    [notice] + docker run --entrypoint=bash -i --rm -v `pwd`:/pwd --workdir /pwd -t 'example_{%id%}_with_setup:latest'
    |    root@93c38006a124:/pwd#
    |
    |", 
    .dir = paste0(temp_dir, "/", id)
  )
})
```
:::

This is useful for interactively debugging issues inside the container. For example, for figuring out whether you need to use `apk`, `apt` or `yum` to install software and to search for the exact name of packages like `libcurl4-openssl-dev`.

## Alternative solutions

There are multiple ways you might try to find a Docker image which contains the right set of dependencies for your component:

* **Browse Docker Hub**: Look a Docker image on Docker Hub or other Docker registries which has the right set of dependencies.
  - This is generally not recommended because it might take a long time to find a pre-existing image with the right set of dependencies
  - Poses a serious security risk. 
* **Write a custom Dockerfile**: You can write a custom Dockerfile to build your own Docker image and store it in a Docker registry, effectively creating a new 'trusted' base image. 
  - Requires manual bookkeeping of which Docker images are used in which components.
  - Not difficult but requires more know-how on how to build custom Docker images.
* **Use Viash setup to build component-specific images**: The methodology described above.
  - Easier to add / change dependencies to one component without breaking another
  - Store images in a centralized container registry

## Behind the screens

### Auto-mount
Any executable built by Viash with a Docker platform will automatically mount the directories of files passed to the executable as arguments. For example, when running:

```bash
./my_executable --input /foo/bar/file.txt --output /dest/path
```

The executable will automatically mount the `/foo/bar` and `/dest` folder to `/viash_automount/foo/bar/` and `/viash_automount/dest` inside the Docker container.

### Auto-chown

By default, files created and modified by a Docker container are owned by **root**. By default, Viash automatically changes the owner of any files defined in the config file to the user running the executable. This behaviour can be overridden by setting the [chown](/reference/config/platforms/DockerPlatform.html#chown) setting to `false` in your config file.

Example with standard Docker:

```bash
docker run -v `pwd`:/pwd bash:4.0 touch /pwd/file.txt
ls -l
```

    -rw-r--r--. 1 root     root         0 Jan 26 16:03 file.txt

Example with a Viash executable:

```bash
/my_executable --output file.txt
```

    -rw-r--r--. 1 myuser   myuser        Jan 26 16:03 file.txt
